Component response rate variation underlies the stability of highly complex finite systems

### 2020-04-14 notebook/SI.Rmd
Add changes for minor revision for Sci Rep

---
title: "Component response rate variation underlies the stability of highly complex finite systems"
author: "A. Bradley Duthie"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA alexander.duthie@stir.ac.uk
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \usepackage{float}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
subtitle: Supplementary Information
biblio-style: apalike
---

>>>>

---
title: "Component response rate variation underlies the stability of highly complex finite systems (Supplementary Information)"
author: "A. Bradley Duthie (alexander.duthie@stir.ac.uk)"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \usepackage{float}
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

### 2020-04-14 notebook/SI.tex
Add changes for minor revision for Sci Rep

\usepackage{microtype}

@@

\usepackage{hyperref}

@@

\hypersetup{unicode=true,
            pdftitle={Component response rate variation underlies the stability of complex systems},
            pdfauthor={A. Bradley Duthie},

@@

@@

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Component response rate variation underlies the stability of complex
systems}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\usepackage{float}
\linenumbers
\bibliographystyle{amnatnat}

@@

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

>>>>

\usepackage[]{microtype}

@@

\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}

@@

\hypersetup{
            pdftitle={Component response rate variation underlies the stability of highly complex finite systems (Supplementary Information)},
            pdfauthor={A. Bradley Duthie (alexander.duthie@stir.ac.uk)},

@@

% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}

@@

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\usepackage{float}
\bibliographystyle{amnatnat}

\title{Component response rate variation underlies the stability of highly
complex finite systems (Supplementary Information)}
\author{A. Bradley Duthie
(\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}})}
\date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA}

@@

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

### 2020-04-14 notebook/cover_letter.Rmd
Add changes for minor revision for Sci Rep

\fontsize{13}{22}
\selectfont

@@

My manuscript addresses a question of broad relevance to the physical, life, and social sciences. I demonstrate that when individual components of complex systems vary in the rate at which they respond to system perturbation, the potential for stability is increased.

In my revision, I have carefully considered the comments of three anonymous reviewers. Addressing these reviewer comments has taken longer than initially anticipated, but I believe that the quality of the manuscript has greatly improved for these additional efforts. I now include new analytical and simulation results that investigate the role of system correlation on stability. I also have performed new simulations for commonly observed network structures, including small-world networks, scale-free networks, and cascade food webs. These new results further clarify the key theory, and further demonstrate its generality across complex systems.

>>>>

\fontsize{15}{22}
\selectfont

@@

My revision now addresses all remaining concerns from the previous round of review.

### 2020-04-14 notebook/ms.tex
Add changes for minor revision for Sci Rep

\textbf{Component response rate (\(\boldsymbol{\gamma}\)) variation}. In
a synthesis of eco-evolutionary feedbacks on community stability, Patel
et al. model a system that includes a vector of potentially changing
species densities (\(\mathbf{n}\)) and a vector of potentially evolving
traits
(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
For any species \(i\) or trait \(j\), change in species density
(\(n_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
of the vectors \(\mathbf{n}\) and \(\mathbf{x}\),

@@

In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
effects of all species densities and trait values on the density of a
species \(i\) and the value of trait \(j\), respectively. Patel et al.
were interested in stability when the evolution of traits was relatively
slow or fast in comparison with the change in species
densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, and
this is modulated in the above by the scalar \(\epsilon\). The value of
\(\epsilon\) thereby determines the timescale separation between ecology
and evolution, with high \(\epsilon\) modelling relatively fast
evolution and low \(\epsilon\) modelling relative slow
evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.

I use the same principle that Patel et al. use to modulate the relative
rate of evolution to modulate rates of component responses for \(S\)
components. Following
May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{32}},
the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),

@@

\textbf{Supplementary Information:} Full tables of stability results for
simulations across different system size (\(S\)) values, ecological
community types, connectance (\(C\)) values, interaction strengths
(\(\sigma\)), and \(\gamma\) distributions are provided as supplementary
material. An additional table also shows results for how feasibility
changes across \(S\). All code and simulation outputs are publicly
available as part of the RandomMatrixStability package on GitHub
(\url{https://github.com/bradduthie/RandomMatrixStability}).

@@

\includegraphics{fig1.pdf}

@@

\includegraphics{fig2.pdf}

@@

\includegraphics{fig3.pdf}

@@

\includegraphics{fig4.pdf}

>>>>

\textbf{Component response rate (\(\boldsymbol{\gamma}\)) variation}. In
a synthesis of eco-evolutionary feedbacks on community stability, Patel
et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}} model a system that includes a vector of potentially changing
species densities (\(\mathbf{n}\)) and a vector of potentially evolving
traits
(\(\mathbf{x}\)).
For any species \(i\) or trait \(j\), change in species density
(\(n_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
of the vectors \(\mathbf{n}\) and \(\mathbf{x}\),

@@

In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
effects of all species densities and trait values on the density of a
species \(i\) and the value of trait \(j\), respectively. Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}
were interested in stability when the evolution of traits was relatively
slow or fast in comparison with the change in species
densities, and
this is modulated in the above by the scalar \(\epsilon\). The value of
\(\epsilon\) thereby determines the timescale separation between ecology
and evolution, with high \(\epsilon\) modelling relatively fast
evolution and low \(\epsilon\) modelling relative slow
evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.

I use the same principle that Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}} use to modulate the relative
rate of evolution to modulate rates of component responses for \(S\)
components. Following
May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{32}},
the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),

@@

@@

@@

@@

@@
